---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

In this part we download, unzip, and finally load the data from a CSV
file:

```{r, echo=TRUE}
zip_filename <- "activity.zip"
## The download command is optional as the file is already in the directory
download.file("https://raw.githubusercontent.com/rdpeng/RepData_PeerAssessment1/master/activity.zip", zip_filename, method = "wget", mode="wb")
f <- unz(description = zip_filename, filename = "activity.csv")
dat <- read.csv(f)
summary(dat)
```

## What is mean total number of steps taken per day?

Mean total number of steps taken per day is:

```{r, echo=TRUE}
aggregated_steps <- aggregate(dat$steps, by=list(dat$date), sum)
steps_per_day <- na.omit(aggregated_steps$x)
hist(steps_per_day)
## summary(steps_per_day)
```

The mean total number of steps taken per day is
`r format(mean(steps_per_day), nsmall=1)`,
while the median total number of steps taken per day is
`r median(steps_per_day)`.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
