---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

In this part we download, unzip, load, and finally filter the data
from a CSV file:

```{r, echo=TRUE}
zip_filename <- "activity.zip"
## The download command is optional as the file is already in the directory
download.file("https://raw.githubusercontent.com/rdpeng/RepData_PeerAssessment1/master/activity.zip", zip_filename, method = "wget", mode="wb")
f <- unz(description = zip_filename, filename = "activity.csv")
dat <- read.csv(f)
dat_filtered <- dat[complete.cases(dat), ]
```

## What is mean total number of steps taken per day?

A histogram of the total number of steps taken each day is as follows:
   
```{r, echo=TRUE}
aggregated_steps <- aggregate(dat_filtered$steps, by=list(dat_filtered$date), sum)
steps_per_day <- aggregated_steps$x
hist(steps_per_day)
```

The mean total number of steps taken per day is
`r format(mean(steps_per_day), nsmall=1)`, while the median total
number of steps taken per day is `r median(steps_per_day)`.


## What is the average daily activity pattern?

A time series plot of the 5-minute interval (x-axis) and the average
number of steps taken, averaged across all days (y-axis):

```{r, echo=TRUE}
steps_ts <- aggregate(dat_filtered$steps, by=list(dat_filtered$interval), mean)
max_i <- which(steps_ts$x == max(steps_ts$x))
head(steps_ts)
plot(steps_ts$Group.1, steps_ts$x, type="l", xlab="5-minute intervals", ylab="Avg. number of steps")
```

A 5-minute interval with index `r max_i` (value
`r dat_filtered$interval[max_i]`), on average across all the days in
the dataset, contains the maximum number of steps
(`r steps_ts$x[max_i]`).

## Imputing missing values

```{r, echo=TRUE}
num_missing_values <- nrow(dat) - nrow(dat_filtered)
```

The total number of missing values in the dataset is `r num_missing_values`.

The following chunk replaces NA values in the original dataset with
the mean of the corresponding 5-minute interval's value and stores the
result in a new dataset:

```{r, echo=TRUE}
## Replace NA values with the mean of the corresponding 5-minute interval's value
dat_modified <- dat
dat_modified[is.na(dat_modified), 1] <- steps_ts$x[match(steps_ts$Group.1, dat_modified$interval)]
## head(dat_modified,100)
## summary(dat_modified)
```

A histogram of the total number of steps taken each day based on the modified data is as follows:

```{r, echo=TRUE}
aggregated_steps_modified <- aggregate(dat_modified$steps, by=list(dat_modified$date), sum)
steps_per_day_modified <- aggregated_steps_modified$x
hist(steps_per_day_modified)
```

The mean total number of steps taken per day in the modified dataset is
`r format(mean(steps_per_day_modified), nsmall=1)`, while the median total
number of steps taken per day is
`r format(median(steps_per_day_modified), nsmall=1)`.

While the mean is the same as in the original dataset, the median is
slightly higher.

## Are there differences in activity patterns between weekdays and weekends?

